Towards a standardized evaluation of imputation methodology

### 2022-03-25 Formatted/Formatted.tex

\documentclass[bimj,fleqn]{w-art}
\usepackage{times}
\usepackage{w-thm}
\usepackage[authoryear]{natbib}
\setlength{\bibsep}{2pt}
\setlength{\bibhang}{2em}
\newcommand{\J}{J\"{o}reskog}
\newcommand{\So}{S\"{o}rbom}
\newcommand{\bcx}{{\bf X}}
\newcommand{\bcy}{{\bf Y}}
\newcommand{\bcz}{{\bf Z}}
\newcommand{\bcu}{{\bf U}}
\newcommand{\bcv}{{\bf V}}
\newcommand{\bcw}{{\bf W}}
\newcommand{\bci}{{\bf I}}
\newcommand{\bch}{{\bf H}}
\newcommand{\bcb}{{\bf B}}
\newcommand{\bcr}{{\bf R}}
\newcommand{\bcm}{{\bf M}}
\newcommand{\bcf}{{\bf F}}
\newcommand{\bcg}{{\bf G}}
\newcommand{\bcs}{{\bf S}}
\newcommand{\bca}{{\bf A}}
\newcommand{\bcd}{{\bf D}}
\newcommand{\bcc}{{\bf C}}
\newcommand{\bce}{{\bf E}}
\newcommand{\ba}{{\bf a}}
\newcommand{\bb}{{\bf b}}
\newcommand{\bc}{{\bf c}}
\newcommand{\bd}{{\bf d}}
\newcommand{\bx}{{\bf x}}
\newcommand{\by}{{\bf y}}
\newcommand{\bz}{{\bf z}}
\newcommand{\bu}{{\bf u}}
\newcommand{\bv}{{\bf v}}
\newcommand{\bh}{{\bf h}}
\newcommand{\bl}{{\bf l}}
\newcommand{\be}{{\bf e}}
\newcommand{\br}{{\bf r}}
\newcommand{\bw}{{\bf w}}
\newcommand{\de}{\stackrel{D}{=}}
\newcommand{\bt}{\bigtriangleup}
\newcommand{\bfequiv}{\mbox{\boldmath $\equiv$}}
\newcommand{\bmu}{\mbox{\boldmath $\mu$}}
\newcommand{\bnu}{\mbox{\boldmath $\nu$}}
\newcommand{\bxi}{\mbox{\boldmath $\xi$}}
\newcommand{\btau}{\mbox{\boldmath $\tau$}}
\newcommand{\bgamma}{\mbox{\boldmath $\Gamma$}}
\newcommand{\bphi}{\mbox{\boldmath $\Phi$}}
\newcommand{\bfphi}{\mbox{\boldmath $\varphi$}}
\newcommand{\bfeta}{\mbox{\boldmath $\eta$}}
\newcommand{\bpi}{\mbox{\boldmath $\Pi$}}
\newcommand{\bequiv}{\mbox{\boldmath $\equiv$}}
\newcommand{\bvarepsilon}{\mbox{\boldmath $\varepsilon$}}
\newcommand{\btriangle}{\mbox{\boldmath $\triangle$}}
\newcommand{\bdelta}{\mbox{\boldmath $\Delta$}}
\newcommand{\beps}{\mbox{\boldmath $\epsilon$}}
\newcommand{\btheta}{\mbox{\boldmath $\theta$}}
\newcommand{\balpha}{\mbox{\boldmath $\alpha$}}
\newcommand{\bsphi}{\mbox{\boldmath $\varphi$}}
\newcommand{\bsig}{\mbox{\boldmath $\sigma$}}
\newcommand{\bfpsi}{\mbox{\boldmath $\psi$}}
\newcommand{\bfdelta}{\mbox{\boldmath $\delta$}}
\newcommand{\bsigma}{{\bf \Sigma}}
\newcommand{\bzero}{{\bf 0}}
\newcommand{\bpsi}{\mbox{\boldmath $\Psi$}}
\newcommand{\bep}{\mbox{\boldmath $\epsilon$}}
\newcommand{\bomega}{\mbox{\boldmath $\Omega$}}
\newcommand{\bfomega}{\mbox{\boldmath $\omega$}}
\newcommand{\blambda}{\mbox{\boldmath $\Lambda$}}
\newcommand{\bflambda}{\mbox{\boldmath $\lambda$}}
\newcommand{\bfsigma}{\mbox{\boldmath $\sigma$}}
\newcommand{\bfpi}{{\mbox{\boldmath $\pi$}}}
\newcommand{\bupsilon}{\mbox{\boldmath $\upsilon$}}
\newcommand{\obs}{{\rm obs}}
\newcommand{\mis}{{\rm mis}}
\theoremstyle{plain}
\newtheorem{criterion}{Criterion}
\theoremstyle{definition}
\newtheorem{condition}[theorem]{Condition}
\usepackage[]{graphicx}
\chardef\bslash=`\\ % p. 424, TeXbook
\newcommand{\ntt}{\normalfont\ttfamily}
\newcommand{\cn}[1]{{\protect\ntt\bslash#1}}
\newcommand{\pkg}[1]{{\protect\ntt#1}}
\let\fn\pkg
\let\env\pkg
\let\opt\pkg
\hfuzz1pc % Don't bother to report overfull boxes if overage is < 1pc
\newcommand{\envert}[1]{\left\lvert#1\right\rvert}
\let\abs=\envert

@@

An obvious problem with design-based simulation is that obtaining a large dataset without missing entries can be very challenging. Most real-world data contains at least one missing entry, for which the true underlying missing data model is by definition unknown. Therefore, the simulator needs to deal with missingness in \textit{some} way before incomplete empirical data can serve as comparative truth in the simulations. It may seem like an intuitive solution to only draw complete cases from the large dataset, which would indeed yield complete samples. However, there may be inherent differences between cases with and cases without any missing values, due to the unknown non-response model. Only sampling complete cases would thus result in a simulation that fails to capture all relevant real-world conditions [which refutes the main reason for using a design-based data generating mechanism in the first place]. Another way to deal with missingness in a design-based simulation is to impute the incomplete dataset once, to obtain a single completed dataset to draw samples from. Unfortunately, this practice may favor the imputation method that was used in this initial imputation step throughout any further evaluations. Nowadays, ... [Add that some simulation studies do not use a `ground truth' at all, they just look for the method with the best predictive performance for a certain completely observed target variable. These studies apply different combinations of imputation and estimation methods on one or more benchmark data sets, and assess the RMSE of the predicted values after imputation and estimation to say something about the comparative performance of the methods. In this context, there is no talk of retrieving the `true' but unobserved values at all. Refer to missing data chapter \citep{liu21}.]

[Design-based model simulation??] Something else to consider is the number of samples that are drawn from the data generating mechanism. If evaluating missing data methodology is the primary focus of the simulation, it may not be necessary to draw data repeatedly. Sampling variation has always been an essential part of the evaluation of multiple imputation methodology. However, in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness, it is not necessary to take sampling variation into account \citep{vink14}. After all, we are interested only in the missing data mechanism and are not considering the noise induced by the sampling mechanism for evaluation in such studies. Note that the conventional pooling rules \citep[cf.][p. 76-77]{rubi87} do not apply for finite population inference. Instead, alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}. [An example from our own experience is that it may be difficult to get true parameter estimates of transformed data (in this case, the mean of a variable with a skew normal distribution). If we wouldn't have used a simulation set-up without sampling variance, it may have led to imprecise reference values of the estimands in our simulation. So, on the one hand, the typical set-up obfuscates the process that we're interested in (i.e., varying the missing data problem over simulation repetitions), while on the other hand, SEs may be incorrect if the sampling variance is not handled correctly in subsequent inferences.] % we hebben al sampling variance tussen de incomplete data sets, extra sampling variance op het complete data level vertroebelt het proces alleen maar. we verplaatsen liever de monte carlo naar het niveau van de missingness.

@@

Evaluating imputation methodology requires a missing data problem to be solved. After establishing a comparative truth from a data generating mechanism, some form of missingness has to be induced. Often, however, reports of simulation studies remain vague about the actual missingness conditions under investigation and, even worse, some authors only report something like
\begin{quote}
\textit{We generated missing data following a missing at random missingness mechanism.}
\end{quote}
This should be considered unacceptable as claims about the validity of the imputation inference heavily depend on the simulated missingness conditions, such as missingness patterns and missingness mechanisms. Missingness patterns concern the location of missing entries across variables in an incomplete dataset, whereas missingness mechanisms describe the relationship between missingness and the values of variables in the data \citep[][p. 8]{litt20}. Under this definition, the missingness pattern also encompasses the proportion of missingness in a given dataset.

% L&R, p. 8: "We find it useful to distinguish the missingness pattern, which describes which values are missing and observed in the data matrix and the missingness mechanism (or mechanisms), which concerns the relationship between missingness and the values of variables in the data matrix".

Even the terminology on missingness generation can be confusing. Does a missingness proportion of 50\% mean that half of the entries in an incomplete dataset are missing, or that half of the rows have at least one missing entry? In this paper, we will henceforth refer to the latter as the proportion of incomplete cases, and keep the term `missingness proportion' restricted to the variable-by-variable interpretation. This distinction, however, is not always clear in the literature. Convoluting the two concepts may lead to incorrect generalizations from simulation studies since the two are rarely equal (e.g., a proportion of 50\% incomplete cases could translate to a missingness proportion of just 25\% in bivariate data, or even 10\% in data with five variables). Moreover, the value of the actual missingness proportion may be diffusing too. Some studies use only 10\% missing data [rephrase to use missingness proportion?] where other studies push the limits to additionally investigate the performance under at least 50\% missing data. This inconsistent display of simulation results may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favorable because of the less stringent simulation conditions. This ultimately may lead to statisticians recommending a less efficient method to applied researchers, thereby limiting the efficiency of the imputation approach and unnecessarily lowering the statistical power.

@@

Other problems with missingness patterns entail the distribution of missing values across cases (as opposed to the proportion over variables). [Add: monotone/univariate vs multivariate, sporadic vs systematic, and univariately generated missing data may not have the expected multivariate patterns. For example, iterative imputation algorithms converge instantly if there is only 1 variable with missingness or monotone pattern, but require iteration in all other situations.] Specific missing data patterns may inadvertently guide which imputation method prevails in performance. Some ad hoc methods are known to yield valid inferences under certain restricted conditions (e.g., list-wise deletion may outperform imputation methods if there is missingness in the outcome variable of the analysis model exclusively), while biasing inferences in any other case. Extrapolating simulation results from such a specific missing data pattern to more intricate empirical missingness patterns could lead to false conclusions.

[Missingness mechanisms: 1) definitions of mechanisms, 2) that MCAR is often ignored, while it's the minimum requirement and reference, and sometimes realistic in practice, 3) that MAR can be generated \textit{spuriously} if the correlation between variables is low, and 4) that MNAR is often ignored, while it may be the most realistic in practice, 5) and finally something about types of M(N)AR (ref dance paper rianne).]

@@

[MAR may be spurious:] With MAR missingness mechanisms, observed relations in the data are used to induce missingness during simulation (e.g. weight is made incomplete based on gender) such that the probability to be missing is the same within groups of cases defined by the observed data (e.g. males are less likely to disclose their weight, but gender is observed). The relations in the observed data may, however, be weak or non-existent. If MAR is induced based on weaker relations in the data, claims for a method's applicability to situations where the missingness is random become less valid. In the most extreme case, MAR is induced from data without multivariate relations. The inferential implications of the missingness would then mimic those of MCAR, even though the missingness is random. In fact, this would amount to one of the special cases under which complete case analysis would be more efficient than multiple imputation \citep[see e.g.][p. 48]{fimd}: the missingness does not depend on the incomplete variable. This property might be useful in practice, but considering it as a condition to evaluate performance under MAR missingness is useless.

@@

[MNAR is often ignored:] Missingness mechanisms that are not random (MNAR) and mechanisms that are considered non-ignorable \citep[see e.g.][]{rubi76}, are generally ignored during evaluation of imputation methodologies, except for those methods that are specifically targeted at non-ignorable applications. Although one cannot definitively verify if the missingness is random--after all, for every MNAR mechanism there is a MAR mechanism with equal fit \citep{molenberghs2008every}--it can be argued that MNAR is the more likely mechanism for real-life missingness scenarios.

[Missingness type is often ignored (duplicated text from solutions section): Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model. This may influence the performance of some (but not necessarily all) imputation routines. For example, inference from hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques. It would be a shame to overlook such results due to the focus on a single MAR mechanism.

@@

To obtain valid imputation inference, the imputation model must capture the essence of the true non-response mechanism \citep{meng94}. The model--if any--that is used to generate the missingness is usually assumed to be random (MAR) or completely random (MCAR).

[All imputation models should be congenial, but this is difficult to test so often ignored even in single imputation endeavor, let alone in simulation studies with hundreds of imputations. Checks of imputation model fit are often omitted. Ignoring congeniality may yield sub-optimal imputations, which may disadvantage certain methods compared to others. Same goes for convergence: FCS is iterative and requires algorithmic convergence, but this is typically evaluated through visual inspection which is unfeasible in simulation studies. Problems with convergence may be even more severe: some methods do not produce imputations at all for certain conditions, while at other times the imputation algorithm is cut off too soon, and performance will be under-estimated.]

@@

[Add segway between evaluation of imputation and actual performance measures.]
The evaluation criteria used to assess imputation performance vary from one developer to another. This is not surprising as people from different fields could have a different focus on the problem at hand. But the choice of performance measure(s) may inadvertently distort statistical properties of the imputed data.

[Add choice of analysis model/estimand here, since performance measure are defined wrt the estimand, e.g., focusing on univariate estimates may favor invalid ad hoc methods.]

Developers often only inquire about the 'accuracy' (i.e. how well can the method reproduce the original data). [Add RMSE at cell level here?] The goal of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete. We are interested in the correct answer to the research question; not in the truth itself. This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should be studied. After all, the 95\% confidence interval should contain the `true' value at least 95 out of 100 times \citep[][p. 591]{neym34}. [So, what could go wrong by focusing on accuracy exclusively? See problems with RMSE of predictions.]

[Add problems with RMSE here. And that if the goal is inference (not prediction) then you need appropriate uncertainty, so m = 1 does not suffice.]

@@

[In general, adhere to ADEMP \citep{morr18}. But there are some differences for the specific case of imputation methodology. All of the parts in ADEMP are used, but a bit scrambled up. Table 1 outlines some of the steps to consider in simulation studies focused on evaluating imputation methodology. The rest of this section is ordered based on these steps.]

\begin{table}[htb]
\begin{center}
\caption{Steps to consider in imputation simulation studies.}
\begin{tabular}{lll}
\hline
Step                    & Aspects under consideration \\
\hline  
1. Set scope            & Aim(s), missing data method(s) to evaluate, number of simulation \\
                        & repetitions \\
2. Obtain truth         & Data generating mechanism(s), estimand(s), sampling variance \\
3. Induce missingness   & Missingness mechanism(s), missing data pattern(s), missingness \\
                        & proportion(s) \\
4. Apply methods        & Imputation model(s), analysis model(s) \\
5. Evaluate imputations & Algorithmic convergence, distributional characteristics, imputation \\
                        & model fit \\
6. Evaluate performance & Performance measures \\
7. Report               & Text, visualization, checklist, simulation script \\
\hline
\end{tabular}
\end{center}
% \footnotesize{Note. Multiple aspects may be varied across simulation conditions (e.g., using a factorial design with several data generating mechanisms and missingness proportions).}
\end{table}

@@

[Make aim clear. Choose scope (type of DGM, target) and imputation methods to evaluate. Ideally, perform fully factorial design, see \citet{morr18}. Think about tolerable level of uncertainty to pick a number of simulation repetitions. Think about general set-up: are the missing data methods all applied to the same incomplete dataset in each simulation repetition, or is a new incomplete set created per method?]

@@

In model-based simulation data are drawn from a known probability distribution, so the theoretical parameters under which the samples are obtained serve as the comparative truth in the simulations. In design-based simulation, data are sampled from a sufficiently large set, so the parameters of this specific set serve as the comparative truth in the simulations.

The design-based approach is often used in situations where a probability distribution is not available, or where real-life data structures are of interest. Applications of design-based simulation are often found in official statistics. A benefit of design-based simulation is the ability to use real-life observed data structures.

For evaluations of model-based simulations, it could be convincing to demonstrate a method's applicability to real-world missing data problems. This can, for example, be achieved by obtaining and imputing a fully observed set of variables, wherein the missingness is mimicked from a similar, incomplete set. Alternatively, an incomplete set could be obtained and truth could be established by removing the incomplete cases from the data. However, the real-world missingness would then be omitted.

The simulator must decide whether there is a need for sampling variance in the simulations scheme. For situations where sampling variance is not of interest, the sampling process can be omitted and a single complete dataset can simply be obtained from the data generating mechanism. The parameters of that single complete set will serve as comparative truth. This process is computationally convenient because only a single complete data set has to be considered during all of the simulations. Use alternative pooling rules for finite population inference \citep{raghunathan2003multiple,vink14}. If sampling variance cannot be omitted from the simulation scheme, draw data repeatedly and apply conventional pooling rules \citep[cf.][p. 76-77]{rubi87}.

@@

Although we limit the focus here to ignorable non-response, the suggested proportions are equally applicable to simulations under non-ignorable non-response.

@@

1. set-up

>>>>

\documentclass[bimj,fleqn]{w-art}
\usepackage{times}
\usepackage{w-thm}
\usepackage[authoryear]{natbib}
% \setlength{\bibsep}{2pt}
% \setlength{\bibhang}{2em}
% \newcommand{\J}{J\"{o}reskog}
% \newcommand{\So}{S\"{o}rbom}
% \newcommand{\bcx}{{\bf X}}
% \newcommand{\bcy}{{\bf Y}}
% \newcommand{\bcz}{{\bf Z}}
% \newcommand{\bcu}{{\bf U}}
% \newcommand{\bcv}{{\bf V}}
% \newcommand{\bcw}{{\bf W}}
% \newcommand{\bci}{{\bf I}}
% \newcommand{\bch}{{\bf H}}
% \newcommand{\bcb}{{\bf B}}
% \newcommand{\bcr}{{\bf R}}
% \newcommand{\bcm}{{\bf M}}
% \newcommand{\bcf}{{\bf F}}
% \newcommand{\bcg}{{\bf G}}
% \newcommand{\bcs}{{\bf S}}
% \newcommand{\bca}{{\bf A}}
% \newcommand{\bcd}{{\bf D}}
% \newcommand{\bcc}{{\bf C}}
% \newcommand{\bce}{{\bf E}}
% \newcommand{\ba}{{\bf a}}
% \newcommand{\bb}{{\bf b}}
% \newcommand{\bc}{{\bf c}}
% \newcommand{\bd}{{\bf d}}
% \newcommand{\bx}{{\bf x}}
% \newcommand{\by}{{\bf y}}
% \newcommand{\bz}{{\bf z}}
% \newcommand{\bu}{{\bf u}}
% \newcommand{\bv}{{\bf v}}
% \newcommand{\bh}{{\bf h}}
% \newcommand{\bl}{{\bf l}}
% \newcommand{\be}{{\bf e}}
% \newcommand{\br}{{\bf r}}
% \newcommand{\bw}{{\bf w}}
% \newcommand{\de}{\stackrel{D}{=}}
% \newcommand{\bt}{\bigtriangleup}
% \newcommand{\bfequiv}{\mbox{\boldmath $\equiv$}}
% \newcommand{\bmu}{\mbox{\boldmath $\mu$}}
% \newcommand{\bnu}{\mbox{\boldmath $\nu$}}
% \newcommand{\bxi}{\mbox{\boldmath $\xi$}}
% \newcommand{\btau}{\mbox{\boldmath $\tau$}}
% \newcommand{\bgamma}{\mbox{\boldmath $\Gamma$}}
% \newcommand{\bphi}{\mbox{\boldmath $\Phi$}}
% \newcommand{\bfphi}{\mbox{\boldmath $\varphi$}}
% \newcommand{\bfeta}{\mbox{\boldmath $\eta$}}
% \newcommand{\bpi}{\mbox{\boldmath $\Pi$}}
% \newcommand{\bequiv}{\mbox{\boldmath $\equiv$}}
% \newcommand{\bvarepsilon}{\mbox{\boldmath $\varepsilon$}}
% \newcommand{\btriangle}{\mbox{\boldmath $\triangle$}}
% \newcommand{\bdelta}{\mbox{\boldmath $\Delta$}}
% \newcommand{\beps}{\mbox{\boldmath $\epsilon$}}
% \newcommand{\btheta}{\mbox{\boldmath $\theta$}}
% \newcommand{\balpha}{\mbox{\boldmath $\alpha$}}
% \newcommand{\bsphi}{\mbox{\boldmath $\varphi$}}
% \newcommand{\bsig}{\mbox{\boldmath $\sigma$}}
% \newcommand{\bfpsi}{\mbox{\boldmath $\psi$}}
% \newcommand{\bfdelta}{\mbox{\boldmath $\delta$}}
% \newcommand{\bsigma}{{\bf \Sigma}}
% \newcommand{\bzero}{{\bf 0}}
% \newcommand{\bpsi}{\mbox{\boldmath $\Psi$}}
% \newcommand{\bep}{\mbox{\boldmath $\epsilon$}}
% \newcommand{\bomega}{\mbox{\boldmath $\Omega$}}
% \newcommand{\bfomega}{\mbox{\boldmath $\omega$}}
% \newcommand{\blambda}{\mbox{\boldmath $\Lambda$}}
% \newcommand{\bflambda}{\mbox{\boldmath $\lambda$}}
% \newcommand{\bfsigma}{\mbox{\boldmath $\sigma$}}
% \newcommand{\bfpi}{{\mbox{\boldmath $\pi$}}}
% \newcommand{\bupsilon}{\mbox{\boldmath $\upsilon$}}
% \newcommand{\obs}{{\rm obs}}
% \newcommand{\mis}{{\rm mis}}
% \theoremstyle{plain}
% \newtheorem{criterion}{Criterion}
% \theoremstyle{definition}
% \newtheorem{condition}[theorem]{Condition}
% \usepackage[]{graphicx}
% \chardef\bslash=`\\ % p. 424, TeXbook
% \newcommand{\ntt}{\normalfont\ttfamily}
% \newcommand{\cn}[1]{{\protect\ntt\bslash#1}}
% \newcommand{\pkg}[1]{{\protect\ntt#1}}
% \let\fn\pkg
% \let\env\pkg
% \let\opt\pkg
% \hfuzz1pc % Don't bother to report overfull boxes if overage is < 1pc
% \newcommand{\envert}[1]{\left\lvert#1\right\rvert}
% \let\abs=\envert

@@

An obvious problem with design-based simulation is that obtaining a large dataset without missing entries can be very challenging. Most real-world data contains at least one missing entry, for which the true underlying missing data model is by definition unknown. Therefore, the simulator needs to deal with missingness in \textit{some} way before incomplete empirical data can serve as comparative truth in the simulations. It may seem like an intuitive solution to only draw complete cases from the large dataset, which would indeed yield complete samples. However, there may be inherent differences between cases with and cases without any missing values, due to the unknown missing data model. Only sampling complete cases from the data set may thus result in samples that fail to capture all relevant real-world conditions [which in turn refutes the main reason for using design-based simulation]. Another way to deal with missingness in a design-based simulation is to impute the incomplete dataset once, to obtain a single completed dataset to draw samples from. Unfortunately, this practice may favor the imputation method that was used in this initial imputation step throughout any further evaluations. [Just to be clear: leaving the missingness as-is and inducing additional missing values to impute doesn't work here, because we wouldn't have a real comparative truth (other than CCA estimates from the large data set, which may be biased).]

After deciding upon a data generating mechanism, another influential--but often overlooked--decision awaits. Simulation studies on missing data methodolology have the unique option to exclude sampling variance from their evaluations, and only use the missing data generation procedure as source of Monte Carlo variation. After all, we are interested in the missingness and are not considering the noise induced by the sampling mechanism for evaluation in such studies. Therefore, it is not necessary to take sampling variation into account in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness \citep{vink14}. A lot of published work does not consider this option, while it could have sharpened inconclusive or obfuscated results. [An example from our own experience is that it may be difficult to get true parameter estimates of transformed data (in this case, the mean of a variable with a skew normal distribution). If we would have used a simulation set-up with sampling variance, it may have led to imprecise reference values of the estimands in our simulation. Table \ref{table:dgm_truth} displays the source of the comparative truth in simulation studies with and without sampling variance.] There is a downside, however, in deviating from the standard simulation workflow by adapting a set-up withoput sampling variance, namely an increased potential for errors in pooling multiple imputation inferences. The conventional pooling rules \citep[cf.][p. 76-77]{rubi87} do not apply for finite population inference. Instead, alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}. If the sampling variance is not handled correctly, subsequent inferences may be invalid and the simulation results cannot be trusted.

@@

Evaluating imputation methodology requires a missing data problem to be solved. After establishing a comparative truth from a data generating mechanism, some form of missingness has to be induced. Often, however, reports of simulation studies remain vague about the actual missingness conditions under investigation and, even worse, some authors only report something like
\begin{quote}
\textit{We generated missing data following a missing at random missingness mechanism.}
\end{quote}
This should be considered unacceptable as claims about the validity of the imputation inference heavily depend on the simulated missingness conditions, such as missingness patterns and missingness mechanisms. Missingness patterns concern the location of missing entries across incomplete data, whereas missingness mechanisms describe the relationship between missing entries and observed data values \citep[][p. 8]{litt20}. [Under this definition, there is a row-wise element to the missingness pattern describing which variables are jointly observed, and a column-wise element encompassing the proportion of missingness in the data.]

[Even the terminology on missingness generation can be confusing. Does a missingness proportion of 50\% mean that half of the entries in an incomplete dataset are missing, or that half of the rows have at least one missing entry?] In this paper, we will henceforth refer to the latter as the proportion of incomplete cases, and keep the term `missingness proportion' restricted to the variable-by-variable interpretation. The distinction between the two concepts, however, is not always clear in the literature, and convoluting the terms may lead to incorrect generalizations because they rarely mean the same (e.g., a proportion of 50\% incomplete cases in bivariate data could translate to a missingness proportion 25\% in both variables, or one completely observed variable and one with 50\% missingness). Moreover, the value of the actual missingness proportion may be diffusing too. Some studies use only 10\% missing data [rephrase to use missingness proportion?] where other studies push the limits to additionally investigate the performance under missingness proportions of at least 50\%. The inconsistent display of simulation conditions may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favorable because of less stringent simulation conditions. This ultimately may lead to statisticians recommending a less efficient method to applied researchers, thereby limiting the efficiency of the imputation approach and unnecessarily lowering the statistical power.

@@

[Other problems with missingness patterns: monotone/univariate vs multivariate, sporadic vs systematic, and univariately generated missing data may not have the expected multivariate patterns. For example, iterative imputation algorithms converge instantly if there is only 1 variable with missingness or monotone pattern, but require iteration in all other situations. And very specific missing data patterns may misguide which imputation method prevails in performance: some ad hoc methods are known to yield valid inferences under certain restricted conditions (e.g., list-wise deletion may outperform imputation methods if there is missingness in the outcome variable of the analysis model exclusively), while biasing inferences in any other case. Extrapolating simulation results from such a specific missing data pattern to more intricate empirical missingness patterns could lead to false conclusions. ]

[Missingness mechanisms paragraphs: 1) definitions of mechanisms (now in table), 2) that MCAR is often ignored, while it's the minimum requirement and reference, and sometimes realistic in practice, 3) that MAR can be generated \textit{spuriously} if the correlation between variables is low,  4) that MNAR is often ignored, while it may be the most realistic in practice, and 5) finally add something about types of M(N)AR (ref Rianne's Dance paper).]

@@

[MAR may be spurious:] With MAR missingness mechanisms, observed relations in the data are used to induce missingness during simulation (e.g. weight is made incomplete based on gender to emulate a situation in which male participants are less likely to disclose their weight, but gender is observed). These relations in the observed data may, however, be weak or non-existent. If MAR is induced based on weaker relations in the data, claims for a method's applicability to situations where the missingness is random become less valid. In the most extreme case, MAR is induced from data without multivariate relations. The inferential implications of the missingness would then mimic those of MCAR, even though the missingness is random. In fact, this would amount to one of the special cases under which complete case analysis would be more efficient than multiple imputation \citep[see e.g.][p. 48]{fimd}: the missingness does not depend on the incomplete variable. This property might be useful in practice, but considering it as a condition to evaluate performance under MAR missingness is useless. %The problem of spurious MAR is generally overlooked in simulation studies

@@

[MNAR is often ignored:] Although one cannot definitively verify if the missingness is random--after all, for every MNAR mechanism there is a MAR mechanism with equal fit \citep{molenberghs2008every}--it can be argued that MNAR is the more likely mechanism for real-life missingness scenarios. Missingness mechanisms that are not random (MNAR) and mechanisms that are considered non-ignorable \citep[see e.g.][]{rubi76}, are generally ignored during evaluation of imputation methodologies, except for those methods that are specifically targeted at non-ignorable applications.

[Missingness type is often ignored (text duplicated from solutions section):] Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model. This may influence the performance of some (but not necessarily all) imputation routines. For example, inference from hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques. It would be a shame to overlook such results due to the focus on a single MAR mechanism.

@@

[All imputation models should be congenial, but this is difficult to test so often ignored even in single imputation endeavor, let alone in simulation studies with hundreds of imputations. Checks of imputation model fit are often omitted. Ignoring congeniality may yield sub-optimal imputations, which may disadvantage certain methods compared to others. Same goes for convergence: FCS is iterative and requires algorithmic convergence, but this is typically evaluated through visual inspection which is unfeasible in simulation studies. Problems with convergence may be even more severe: some methods do not produce imputations at all for certain conditions, while at other times the imputation algorithm is cut off too soon, and performance will be under-estimated.] To obtain valid imputation inference, the imputation model must capture the essence of the true non-response mechanism \citep{meng94}. % The model--if any--that is used to generate the missingness is usually assumed to be random (MAR) or completely random (MCAR).

@@

[Add segway between evaluation of imputation and actual performance measures.]
The evaluation criteria used to assess imputation performance vary from one developer to another. This is not surprising as people from different fields could have a different focus on the problem at hand. [Add that simulation aim/target determines which performance measures are relevant to consider, and that the choice of analysis model/estimand can influence the values of these metrics (e.g., focusing on univariate estimates only may favor ad hoc methods).]

The choice of performance measures may inadvertently distort statistical properties of the imputed data. Developers often only inquire about the `accuracy' (i.e. how well can the method reproduce the original data). [Add RMSE at cell level here?]

If the goal is inference (not prediction) then you need appropriate uncertainty, for which a single imputation does not suffice. The goal of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete. We are interested in the correct answer to the research question; not in the truth itself. This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should be studied. After all, the 95\% confidence interval should contain the `true' value at least 95 out of 100 times \citep[][p. 591]{neym34}. [So, what could go wrong by focusing on accuracy exclusively? Add problems with RMSE of predictions.]

@@

[In general, adhere to ADEMP \citep{morr18}. But there are some differences for the specific case of evaluating imputation methodology. All of the parts in ADEMP are used, but a bit scrambled up. Table \ref{table:steps} outlines some of the steps to consider for the evaluation of imputation methodology. The rest of this section is ordered based on these steps.]

\begin{table}[htb]
\begin{center}
\caption{Steps to consider in imputation simulation studies.}
\label{table:steps}
\begin{tabular}{lll}
\hline
Step                    & Aspects under consideration \\
\hline  
1. Set scope            & Aim(s), missing data method(s) to evaluate, precision level, number \\
                        & of simulation repetitions \\
2. Obtain truth         & Data generating mechanism(s), estimand(s), sampling variance \\
3. Induce missingness   & Missingness mechanism(s), missing data pattern(s), missingness \\
                        & proportion(s) \\
4. Apply methods        & Imputation model(s), analysis model(s) \\
5. Evaluate imputations & Algorithmic convergence, distributional characteristics, imputation \\
                        & model fit \\
6. Evaluate performance & Performance measures \\
7. Report               & Text, visualization, checklist, simulation script \\
\hline
\end{tabular}
\end{center}
% \footnotesize{Note. Multiple aspects may be varied across simulation conditions (e.g., using a factorial design with several data generating mechanisms and missingness proportions).}
\end{table}

@@

[Substeps for the simulator include: Define aim. Choose scope (e.g. data source, target) and imputation methods to evaluate. Ideally, perform fully factorial design, see \citet{morr18}. Set tolerable level of uncertainty and use this to pick a number of simulation repetitions. Choose general set-up/simulation flow: are the missing data methods all applied to the same incomplete dataset in each simulation repetition, or is a new incomplete set created per method?]

[Add that some simulation studies do not use a `ground truth' at all, they just look for the method with the best predictive performance for a certain completely observed target variable. These studies apply different combinations of imputation and estimation methods on one or more benchmark data sets, and assess the RMSE of the predicted values after imputation and estimation to say something about the comparative performance of the methods. Refer to missing data chapter \citep{liu21}. In this context, there is no talk of retrieving the `true' but unobserved values at all. The DGM is by definition design-based and there is no sampling variance. The missingness generation step is skipped and the only performance measure that can be used between methods is RMSE and maybe comparison with repect to CCA in terms of bias.]

@@

[Choose sampling variance:] The simulator must decide whether there is a need for sampling variance in the simulations scheme. For situations where sampling variance is not of interest, the sampling process can be omitted and a single complete dataset can simply be obtained from the data generating mechanism. The parameters of that single complete set will serve as comparative truth. This process is computationally convenient because only a single complete data set has to be considered during all of the simulations. Be carefull to use alternative pooling rules for finite population inference \citep{raghunathan2003multiple,vink14}. If sampling variance cannot be omitted from the simulation scheme, draw data repeatedly and apply conventional pooling rules \citep[cf.][p. 76-77]{rubi87}.

[Choose DGM:] The design-based approach is often used in situations where a probability distribution is not available, or where real-life data structures are of interest. Applications of design-based simulation are often found in official statistics. A benefit of design-based simulation is the ability to use real-life observed data structures. For evaluations of model-based simulations, it could be convincing to demonstrate a method's applicability to real-world missing data problems. This can, for example, be achieved by obtaining and imputing a fully observed set of variables, wherein the missingness is mimicked from a similar, incomplete set. Alternatively, an incomplete set could be obtained and truth could be established by removing the incomplete cases from the data. However, the real-world missingness would then be omitted.

[Obtain truth: In model-based simulation data are drawn from a known probability distribution, so the theoretical parameters under which the samples are obtained serve as the comparative truth in the simulations. In design-based simulation, data are sampled from a sufficiently large set, so the parameters of this specific set serve as the comparative truth in the simulations. But this is not a complete picture, see Table \ref{dgm_truth}.]

\begin{table}[htb]
\begin{center}
\caption{Comparative truth in model-based and design-based simulation.}
\label{table:dgm_truth}
\begin{tabular}{lll}
\hline
               & With sampling variance      & Without sampling variance \\
               & (multiple samples drawn)    & (one sample drawn) \\
\hline  
Model-based    & Truth from model            & Truth from sample \\
Design-based   & Truth from large data set   & Truth from sample \\
\hline
\end{tabular}
\end{center}
\end{table}

@@

@@

1. Set-up
